Interspecies comparisons of Mg/Ca ratios in limpet shells

### 2024-06-23 Manuscript.qmd
added citation
Modern shells

The analysed specimens, their origin and studies with background information can be found in Table 1.

@@

: Table 1 Overview of the modern and archaeological limpet specimens analysed in this study {#tab:specimens}

>>>>

The analysed specimens, their origin and respective studies with research background information can be found in [Table 1](#Table_1). Here we will briefly summarise their contextual information, which can be accessed in more detail at the respective studies [@Nicastro2020-ih; @Surge2012-ba; @Graniero2017-io]. While those studies also included other specimens, their accessibility or state of preservation did not all lend them to be re-analysed.

::: {#Table_1}

@@

: Overview of the modern and archaeological limpet specimens analysed in this study.
:::

```{r SST Data}

# Read the CSV files, skipping the comments
Westray_SST <- read_csv(here("data","Westray-SST.csv"), comment = "#")
TdF_SST <- read_csv(here("data","Tierra_del_Fuego-SST.csv"), comment = "#")

# Combine the datasets with new names and add year and month columns
SST_df <- bind_rows(Westray_SST %>% mutate(dataset = "Westray - Orkney"),
                         TdF_SST %>% mutate(dataset = "Cambaceres Bay - Tierra del Fuego")) %>%
  mutate(analysed_sst = analysed_sst - 273.15,
         time = as_datetime(time),
         year = year(time),
         month = month(time, label = TRUE)) %>%
  group_by(dataset, year, month) %>%
  summarise(mean_sst = mean(analysed_sst, na.rm = TRUE),
            sd_sst = sd(analysed_sst, na.rm = TRUE))

WestrayMin <- SST_df %>% filter(dataset=="Westray - Orkney") %>% 
  pull(mean_sst) %>% min() %>% round(1)
WestrayMax <- SST_df %>% filter(dataset=="Westray - Orkney") %>% 
  pull(mean_sst) %>% max() %>% round(1)

CamBayMin <- SST_df %>% filter(dataset=="Cambaceres Bay - Tierra del Fuego") %>% 
  pull(mean_sst) %>% min() %>% round(1)
CamBayMax <- SST_df %>% filter(dataset=="Cambaceres Bay - Tierra del Fuego") %>% 
  pull(mean_sst) %>% max() %>% round(1)

There are two main locations that we sourced specimens from: the Beagle Channel in Tierra del Fuego (Argentina) and the island of West Ray in Orkney (UK). Four specimens were collected from the Beagle Channel in Cambaceres Bay. The area receives around 570 mm precipitation but the channel waters are predominantly influenced by marine currents and in the period of October 2015 to October 2016 stayed around 30.7 ± 0.7 psu [@Nicastro2020-ih]. SSTs range from `r CamBayMin`ºC in August to `r CamBayMax`ºC in February.

On Westray one modern shell (ORK-LT5) was collected in Rack Wick Bay in August 2009. Westray lies about 70 km north of the Scottish mainland and experiences virtually no freshwater input with salinity values of 34.0 – 34.5 psu. The annual SST ranges from `r WestrayMin`ºC in March to `r WestrayMax`ºC in August (@fig-SSTs).

```{r SST plot}
#| fig-cap: SST values averaged from  years. SST values were sourced from [@europeanunion-copernicusmarineservice2015]
#| fig-height: 10
#| fig-width: 10
#| label: fig-SSTs

# Plot the data
SST_df %>% 
    group_by(dataset, month) %>%
  summarise(sd_sst = sd(mean_sst, na.rm = TRUE),
            mean_sst = mean(mean_sst, na.rm = TRUE)) %>%
ggplot(aes(x = month, y = mean_sst, colour = dataset,group=dataset)) +
  geom_ribbon(aes(ymin = mean_sst - sd_sst, ymax = mean_sst + sd_sst,fill=dataset), alpha = 0.2,col="white") +
  geom_path() +
  scale_colour_manual(values = c( "#007f7f", "#EC7063")) +
scale_fill_manual(values = c( "#007f7f", "#EC7063"))+
  labs(title = "Monthly Sea Surface Temperature",
       x = "Month",
       y = "Mean Analysed SST (°C)") +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
    legend.position = c(0.1, 0.9))+  
  scale_y_continuous(breaks = c(5, 10, 15),limits = c(3,15)) +
  labs(colour="",fill="")

### 2024-06-23 Manuscript.tex
Modern shells

The analysed specimens, their origin and respective studies with
research background information can be found in Table 1. Here we will
briefly summarise their contextual information, which can be accessed in
more detail at the respective studies
\citep{Nicastro2020-ih, Surge2012-ba, Graniero2017-io}. While those
studies also included other specimens, their accessibility or state of
preservation did not all lend them to be re-analysed.

There are two main locations that we sourced specimens from: the island
of West Ray in Orkney (UK) and the Beagle Channel in Tierra del Fuego
(Argentina). A modern shell on West Ray was collected in Rack Wick Bay
in August 2009, which experiences virtually no freshwater input. The
temperature

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{Table 1 Overview of the modern and archaeological limpet
specimens analysed in this study \{\#tab:specimens\}}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Modern & \citep{Nicastro2020-ih} & \emph{N. deaureata} & Cambaceres Bay
- Tierra del Fuego (AR) & ND-1016-3 & Stable oxygen and carbon isotope
analysis \\
& & & & ND-1016-4 & \\
& & \emph{N. magellanica} & & NM-1016-1 & \\
& & & & NM-1016-3 & \\
& \citep{Graniero2017-io} & \emph{P. vulgata} & Rack Wick Bay - Westray
(UK) & ORK-LT5 & Stable oxygen and carbon isotope analysis; Mg, Li, Sr,
Ca,. \\
Archaeological & \citep{Surge2012-ba} & \emph{P. vulgata} & Rack Wick
Bay - Westray (UK) & QG1-7188-1 & Stable oxygen and carbon isotope
analysis \\
& & & & QG1-7189-2 & \\
& & & & QG2-1061-1 & \\
& & & & QG2-1064-1 & \\
& & & & QG2-7180-1 & \\
& & & & QG2-7180-2 & \\
& & & & QG1-7246-1 & \\
& This study & \emph{N. magellanica} & Heshkaia 35 in Moat Bay - Tierra
del Fuego (AR) & Heshkaia\_35-1 & \emph{none} \\
& & \emph{N. deaureata} & & Heshkaia\_35-2 & \\
& & \emph{N. deaureata} & & Heshkaia\_35-3 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-4 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-5 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-6 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-7 & \\
& & \emph{N. deaureata} & & Heshkaia\_35-8 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-9 & \\
\end{longtable}

{\centering \includegraphics{Manuscript_files/figure-pdf/SST Data-1.pdf}

@@

\caption{SST values averaged over the last 20 years. SST values were
sourced from \citep{europeanunion-copernicusmarineservice2015}}

>>>>

The analysed specimens, their origin and respective studies with
research background information can be found in \hyperref[Table_1]{Table
1}. Here we will briefly summarise their contextual information, which
can be accessed in more detail at the respective studies
\citep{Nicastro2020-ih, Surge2012-ba, Graniero2017-io}. While those
studies also included other specimens, their accessibility or state of
preservation did not all lend them to be re-analysed.

\phantomsection\label{Table_1}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1667}}@{}}
\caption{Overview of the modern and archaeological limpet specimens
analysed in this study.}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Context
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Study
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Species
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Location
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Sample ID
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Previous analyses
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Modern & \citep{Nicastro2020-ih} & \emph{N. deaureata} & Cambaceres Bay
- Tierra del Fuego (AR) & ND-1016-3 & Stable oxygen and carbon isotope
analysis \\
& & & & ND-1016-4 & \\
& & \emph{N. magellanica} & & NM-1016-1 & \\
& & & & NM-1016-3 & \\
& \citep{Graniero2017-io} & \emph{P. vulgata} & Rack Wick Bay - Westray
(UK) & ORK-LT5 & Stable oxygen and carbon isotope analysis; Mg, Li, Sr,
Ca,. \\
Archaeological & \citep{Surge2012-ba} & \emph{P. vulgata} & Rack Wick
Bay - Westray (UK) & QG1-7188-1 & Stable oxygen and carbon isotope
analysis \\
& & & & QG1-7189-2 & \\
& & & & QG2-1061-1 & \\
& & & & QG2-1064-1 & \\
& & & & QG2-7180-1 & \\
& & & & QG2-7180-2 & \\
& & & & QG1-7246-1 & \\
& This study & \emph{N. magellanica} & Heshkaia 35 in Moat Bay - Tierra
del Fuego (AR) & Heshkaia\_35-1 & \emph{none} \\
& & \emph{N. deaureata} & & Heshkaia\_35-2 & \\
& & \emph{N. deaureata} & & Heshkaia\_35-3 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-4 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-5 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-6 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-7 & \\
& & \emph{N. deaureata} & & Heshkaia\_35-8 & \\
& & \emph{N. magellanica} & & Heshkaia\_35-9 & \\
\end{longtable}

There are two main locations that we sourced specimens from: the Beagle
Channel in Tierra del Fuego (Argentina) and the island of West Ray in
Orkney (UK). Four specimens were collected from the Beagle Channel in
Cambaceres Bay. The area receives around 570 mm precipitation but the
channel waters are predominantly influenced by marine currents stayed
around 30.7 ± 0.7 psu \citep{Nicastro2020-ih}. SSTs range from 4.7ºC in
August to 10ºC in February.

On Westray one modern shell (ORK-LT5) was collected in Rack Wick Bay in
August 2009. Westray lies about 70 km north of the Scottish mainland and
experiences virtually no freshwater input. The annual SST ranges from
6.3ºC in March to 13.8ºC in August (Figure~\ref{fig-SSTs}).

\begin{figure}

\centering{

\includegraphics{Manuscript_files/figure-pdf/fig-SSTs-1.pdf}

@@

\caption{\label{fig-SSTs}SST values averaged from years. SST values were
sourced from \citep{europeanunion-copernicusmarineservice2015}}
